---
title: "Nykaa Marketing Analytics"
author: "Isha Tanwar"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: cosmo
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(lubridate)
library(janitor)
library(kableExtra)
```

## Project Overview

This project analyzes simulated data from Nykaa, a direct-to-consumer beauty and wellness brand, to evaluate the effectiveness of marketing spend, customer behavior, and product ratings.

Objectives:

1. Understand the impact of marketing spend on orders

2. Identify top-performing customers

3. Analyze correlation between product ratings and sales

4. Evaluate conversion rates and channel performance

## Data Import and Cleaning
```{r}
setwd('C:/Users/sujal/OneDrive/Desktop')

# Load datasets
orders <- read.csv("nykaa_orders.csv") %>% clean_names()
customers <- read_csv("nykaa_customers.csv") %>% clean_names()
products <- read_csv("nykaa_products.csv") %>% clean_names()
marketing <- read_csv("nykaa_marketing.csv") %>% clean_names()
reviews <- read_csv("nykaa_reviews.csv") %>% clean_names()

# Convert dates
orders <- orders %>%
  mutate(order_date = dmy(order_date),
         order_month = floor_date(order_date,"month"))

marketing <- marketing %>%
  mutate(marketing_month = ym(month))
```

## Data Integration
```{r}
# Combine all data
order_details <- orders %>%
  left_join(customers, by = "customer_id") %>%
  left_join(products, by = "product_id") %>%
  left_join(reviews, by = "product_id")
```

## Top 10 Customers by Order Volume
```{r}
top_customers <- orders %>%
  group_by(customer_id) %>%
  summarise(TotalOrders = n()) %>%
  arrange(desc(TotalOrders)) %>%
  slice_head(n=10)

top_customers_plot <- ggplot(top_customers,mapping = aes
                             (x = reorder(customer_id,-TotalOrders), 
                                   y = TotalOrders)) +
  geom_col(fill = "#519484") +
  labs(title = "Top 10 Customers by Order Volume", x = "Customer ID", y = "No. of Orders") +
  theme(panel.background = element_blank())
top_customers_plot
```

## Monthly Trend: Marketing Spend vs. Orders
```{r}
monthly_marketing <- marketing %>%
  group_by(marketing_month) %>%
  summarise(total_spend = sum(spend, na.rm = TRUE),
            total_conversions = sum(conversions, na.rm = TRUE))
monthly_orders <- orders %>%
  group_by(order_month) %>%
  summarise(TotalOrders = n())

trend_data <- monthly_marketing %>%
  left_join(monthly_orders, by=c("marketing_month" = "order_month"))

trend_data_graph <- ggplot(trend_data, mapping = aes(x = marketing_month)) +
  geom_line(aes(y = total_spend, colour = "Marketing Spend")) +
  geom_point(aes(y = total_spend, colour = "Marketing Spend")) +
  geom_line(aes(y = TotalOrders*100000, colour = "Total Orders")) +
  geom_point(aes(y = TotalOrders*100000, colour = "Total Orders")) +
  scale_y_continuous(name = "Marketing Spend",sec.axis = 
                       sec_axis(~./100000,name = "Total Orders")) +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 month") +
  theme(panel.background = element_blank(), 
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Monthly Trend: Marketing Spend vs Orders", x = "Marketing Month", colour = "Legend") 
trend_data_graph
```

## Product Ratings vs. Sales
```{r}
product_sales <- orders %>%
  group_by(product_id) %>%
  summarise(unit_sales = n())

product_ratings <- reviews %>%
  group_by(product_id) %>%
  summarise(avg_rating = mean(rating, na.rm = TRUE))
 
ratings_vs_sales <- product_sales %>%
  left_join(product_ratings, by = "product_id")

ratings_vs_sales_graph <- ratings_vs_sales %>%
  ggplot(aes(x = avg_rating, y = unit_sales)) +
  geom_point(color = "#FF69B4") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = "Do Higher-Rated Products Sell More?", 
       x = "Average Rating", y = "Units Sold") +
  theme(panel.background = element_blank())
ratings_vs_sales_graph
```

## Monthly Conversion Rate
```{r}
trend_data <- trend_data %>%
  mutate(conversion_rate = total_conversions/total_spend *100)
trend_data_plot <- ggplot(trend_data, aes(x = marketing_month, y = conversion_rate)) +
  geom_line(color = "#20B2AA") +
  geom_point(color = "#20B2AA") +
  labs(title = "Monthly Conversion Rate from Marketing Campaigns",
       x = "Month", y = "Conversion Rate") +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 month") +
  theme(panel.background = element_blank(), 
        axis.text.x = element_text(angle = 90, hjust = 1))
trend_data_plot
```

## Cost per Conversion by Channel
```{r}
channel_perf <- marketing %>%
  group_by(channel) %>%
  summarise(spend = sum(spend, na.rm=TRUE),
            conversions = sum(conversions, na.rm=TRUE)) %>%
  mutate(cpc = spend/conversions) #costperconversion

channel_perf_plot <- channel_perf %>%
  ggplot(aes(x = reorder(channel,-cpc), y = cpc)) +
  geom_col(fill = "#845194") +
  coord_flip() +
  labs(title = "Cost per Conversion by Channel", x = "Channel", y = "₹ per Conversion") +
  theme(panel.background = element_blank())
channel_perf_plot
```

## Overall KPIs Table
```{r}
kpi_table <- trend_data %>% 
  summarise(`Total Spend (₹M)`      = round(sum(total_spend)/1e6, 2),
            `Total Conversions`     = sum(total_conversions),
            `Total Orders`          = sum(TotalOrders),
            `Overall Conversion %`  = round(sum(total_conversions)/sum(total_spend)*100, 
                                            1))
kpi_table %>% 
  kbl(caption = "Overall Marketing & Sales KPIs") %>% 
  kable_styling(full_width = FALSE, position = "left")
```

## Summary Insights
- Email and Google Ads show the lowest cost per conversion.

- No clear positive correlation found between ratings and units sold.

- Conversion rates fluctuate but peak with increased spend.

- A few customers are contributing to a majority of the orders.

## Conclusion
This analysis helped identify the relationship between marketing investments and sales outcomes. These insights can support better budget allocation, channel selection, and customer retention strategies for D2C brands like Nykaa.











